---
title: "p8105_hw2_ms6360"
author: "Maisie Sun"
date: "2022-10-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

```{r data _import}
NYC_transit = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, starts_with('route'), entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES",TRUE,FALSE))
```

* The dataset contains variables station names, station latitude, different routes (1 to 11), entry, vending, entrance times and ADA. 

* There are `r ncol(NYC_transit)` columns (variables) and `r nrow(NYC_transit)` rows (observations). 

* The data is not tidy and can be improved by having routes as cell number under a variable called "Routes" instead of having different routes as its own variables.

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the `distinct` function may be useful here.

```{r distinct_stations}
NYC_transit %>%
  select(line, station_name) %>%
  distinct
```

* There are `r nrow(distinct(NYC_transit, line, station_name))` distinct stations.

How many stations are ADA compliant?

```{r ADA_compliant_stations}
NYC_transit %>%
  filter(ada == TRUE) %>%
  select(line, station_name) %>%
  distinct
```

* There are 84 stations that are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?

```{r proportion_stations_vending}
NYC_transit %>%
  filter(vending == "NO") %>%
  pull(entry) %>%
  mean
```

* 0.377 of station entrances without vending allow entrance.

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r station_A_trains}
# change column type from double to character
NYC_transit$route8 <- as.character(NYC_transit$route8)  
NYC_transit$route9 <- as.character(NYC_transit$route9)  
NYC_transit$route10 <- as.character(NYC_transit$route10)  
NYC_transit$route11 <- as.character(NYC_transit$route11)  

NYC_transit %>%
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>%
  filter(route == "A") %>%
  select(line, station_name) %>%
  distinct


NYC_transit %>%
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>%
  filter(route == "A", ada == "TRUE") %>%
  select(line, station_name) %>%
  distinct
```

* There are 60 distinct stations on line A and 17 of them are ADA compliant.

# Problem 2

```{r excel_import_joining}
#import Mr. Trash Wheel
trash_wheel = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:O535",
  skip = 1) %>%
janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls, 0)))
  trash_wheel$name <- "Mr.Trash Wheel"

#import Professor Trash Wheel
prof_trash_wheel = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Professor Trash Wheel",
  range = "A2:N117",
  skip = 1) %>%
janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls, 0))) %>%
  add_column(name = "Professor Trash Wheel") %>%
  mutate(dumpster = as.character(dumpster))
  
#fully join datasets
joint_trash_wheel = 
  full_join(trash_wheel, prof_trash_wheel)
```

For Mr. Trash Wheel dataset, there are `r ncol(trash_wheel)` columns (variables) and `r nrow(trash_wheel)` rows (observations). As for Professor Trash Wheel, there are `r ncol(prof_trash_wheel)` columns (variables) and `r nrow(prof_trash_wheel)` rows (observations). Once combined, there are `r ncol(joint_trash_wheel)` columns (variables) and `r nrow(joint_trash_wheel)` rows (observations). Variables in this dataset include `r colnames(joint_trash_wheel)`. Within these variables, the weight of trash collected in tons, number of trash collected (plastic bottles, polystyrene, cigarette butts, grocery bags, glass bottles, chip bags, and sports balls), and time of trash collection were detailed.

The total weight of trash collected by Professor Trash Wheel is `r sum(prof_trash_wheel$weight_tons)`. The total number of sports balls collected by Mr. Trash Wheel in 2020 is `r  trash_wheel %>% filter(year == "2020") %>% pull(sports_balls) %>% sum()`.

# Problem 3

```{r pols_months}
pols_months = 
  read_csv("./data/data3/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep = "-") %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  pivot_longer( 
    cols = starts_with("prez_"),
    names_to = "president",
    names_prefix = "prez_",
    values_to = "number") %>%
  select(-day) %>%
  mutate(year = as.double(year))
```

```{r snp}
snp = 
  read_csv("./data/data3/snp.csv") %>%
  janitor::clean_names()
  
snp_after_01 <- filter(snp, row_number() < 176) %>%
  mutate(date = as.Date(date,"%y/%m/%d")) %>%
  separate(date, into = c("year", "month", "day"), sep = "-")

snp_before_01 <- filter(snp, row_number() > 175) %>%
  mutate(date = as.Date(date,"%m/%d/%y")) %>%
  separate(date, into = c("year", "month", "day"), sep = "-")

snp = 
  full_join(snp_after_01, snp_before_01) %>%
  mutate(month = month.abb[as.numeric(month)]) %>%
  select(-day) %>%
  mutate(year = as.double(year))
```

```{r unemployment}
unemployment = 
  read_csv("./data/data3/unemployment.csv") %>%
  janitor::clean_names() %>%
   pivot_longer( 
     cols = jan:dec,
     names_to = "month",
     values_to = "percent_unemployment")

pols_snp = 
  full_join(pols_months, snp)
pols_snp_unemployment = 
  full_join(pols_snp, unemployment)
```

In the `pols_month` dataset, there are `r nrow(pols_months)` rows (observations) and `r ncol(pols_months)` columns (variables). The range of years is (`r range(pols_months$year)`). Variables in this dataset (`r colnames(pols_months)`) include the number of republican and democratic presidents, governors, senators, and representatives.

For the `snp` dataset, there are `r nrow(snp)` rows (observations) and `r ncol(snp)` columns (variables). The range of years is (`r range(snp$year)`). Variables in this dataset (`r colnames(snp)`) include the closing values of the S&P stock index on the associated month

Finally for the `unemployment` dataset, there are `r nrow(unemployment)` rows (observations) and `r ncol(unemployment)` columns (variables).The range of years is (`r range(unemployment$year)`). Variables in this dataset (`r colnames(unemployment)`) include represents the percentage of unemployment associated with the month.