---
title: "p8105_hw2_ms6360"
author: "Maisie Sun"
date: "2022-10-01"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
```

# Problem 1

```{r data _import}
NYC_transit = 
  read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line, station_name, station_latitude, starts_with('route'), entry, vending, entrance_type, ada) %>% 
  mutate(entry = ifelse(entry == "YES",TRUE,FALSE))

NYC_transit

```

* The dataset contains variables station names, station latitude, different routes (1 to 11), entry, vending, entrance times and ADA. 

* There are `r ncol(NYC_transit)` columns (variables) and `r nrow(NYC_transit)` rows (observations). 

* The data is not tidy and can be improved by having routes as cell number under a variable called "Routes" instead of having different routes as its own variables.

Answer the following questions using these data:

How many distinct stations are there? Note that stations are identified both by name and by line (e.g. 125th St 8th Avenue; 125st Broadway; 125st Lenox); the `distinct` function may be useful here.

```{r distinct_stations}
NYC_transit %>%
  select(line, station_name) %>%
  distinct
```

* There are `r nrow(distinct(NYC_transit, line, station_name))` distinct stations.

How many stations are ADA compliant?

```{r ADA_compliant_stations}
NYC_transit %>%
  filter(ada == TRUE) %>%
  select(line, station_name) %>%
  distinct
```

* There are 84 stations that are ADA compliant.

What proportion of station entrances / exits without vending allow entrance?

```{r proportion_stations_vending}
NYC_transit %>%
  filter(vending == "NO") %>%
  pull(entry) %>%
  mean
```

* 0.377 of station entrances without vending allow entrance.

Reformat data so that route number and route name are distinct variables. How many distinct stations serve the A train? Of the stations that serve the A train, how many are ADA compliant?

```{r station_A_trains}
NYC_transit

# specify the columns I want to change
NYC_transit$route8 <- as.character(NYC_transit$route8)  
NYC_transit$route9 <- as.character(NYC_transit$route9)  
NYC_transit$route10 <- as.character(NYC_transit$route10)  
NYC_transit$route11 <- as.character(NYC_transit$route11)  

NYC_transit %>%
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>%
  filter(route == "A") %>%
  select(line, station_name) %>%
  distinct


NYC_transit %>%
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>%
  filter(route == "A", ada == "TRUE") %>%
  select(line, station_name) %>%
  distinct
```

* There are 60 distinct stations on line A and 17 of them are ADA compliant.

# Problem 2

```{r excel_import}
trash_wheel = 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx",
  sheet = "Mr. Trash Wheel",
  range = "A2:N535",
  skip = 1) %>%
janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = as.integer(round(sports_balls, 0)))

```

